>

### 2024-08-10 docs/article.qmd
Revise the introduction based on suggestion for the ISPOR abstract

funding: "The author(s) received no specific funding for this work."
keywords:
  - Graph Theory
  - Prescription Registry
  - Polypharmacy
  - Mental Health   
  - Public Health Surveillance

@@

@@

In psychiatry, polypharmacy is defined as a patient with at least two concurrent psychiatric medications [@Shrivastava2013]. With a prevalence between 13-90%, psychiatric polypharmacy manifests in five categories, namely same-class, multi-class, adjunctive, augmentation, and total polypharmacy. Same-class polypharmacy is the use of multiple medications from the same class. Multi-class polypharmacy is the use of multiple medications from different classes indicated for the same symptom cluster. Adjunctive polypharmacy is the use of additional medications to treat side effects due to other medications. Augmentation polypharmacy is the use of full-dose and sub-dose medications from a different class for the same symptom cluster. Finally, total polypharmacy is the total count of all medications used by a patient, both psychopharmaca and other classes alike [@Shrivastava2013]. For policymakers, differences in defining polypharmacy pose a challenge to design a surveillance system [@Sirois2016].

Surveilling polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy surveillance system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To reduce the complexity, graph theory offers a complementary approach by analyzing polypharmacy as a drug prescription network (DPN). Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

A DPN is constructed as a matrix which rows to columns represent pairs of medication being taken by a patient in a day. Since there is no limit for the number of rows and columns, a DPN can contain virtually every possible combination of polypharmacy regimens. A DPN is generalizable as a population aggregate by coalescing all single-patient DPN through matrix addition [@Cavallo2012]. Constructing a DPN allows examining co-prescription dynamics in the population as graph measures, which is also attributable to each medication. One of such measures is graph centrality, where we calculate scores representing the number of co-medications, number of connecting paths, average distance, and relative importance [@Askar2021]. The relative importance score is of particular interest to address concerns raised by @Sirois2016 and @Delara2022 regarding standardized quantitative indicator of polypharmacy. Relative importance, most commonly measured by eigenvector centrality, represents the connection to vital medications. A high eigenvector centrality in a DPN indicates more prescriptions with highly-prescribed medications. Following up @Cavallo2012 and @Bazzoni2015, we hypothesize that eigenvector centrality as a measure of relative importance is attributable to indicate polypharmacy. Our investigation showcases eigenvector centrality as a quantitative indicator of polypharmacy to promote a systematic public health surveillance system.

>>>>

funding: "The author(s) received no specific funding for this work."
keywords:
  - Graph Theory
  - Prescription Registry
  - Polypharmacy
  - Mental Health   
  - Public Health Monitoring

@@

::: {.content-hidden}

@@

:::

The relational characteristics in health research variables has fueled the growing interest of implementing network analysis, making it an indispensable research framework in medicine and epidemiology [@Luke2007]. Past implementations of network analysis in public health mostly focused on transmission network, social network, and organizational network. Only recently, @Cavallo2012 demonstrated a novel framework to describe medication uses on a population level, i.e. drug prescription network (DPN), a term which later is coined by @Bazzoni2015. Network analysis offers an approach to capture the relationship among building blocks of a complex system. From a health system standpoint, polypharmacy is a prescription of multiple drug combinations which can be further evaluated as a network [@Miglio2021]. DPN is a complex and dynamic system that may change overtime, specifically also in drug prescribed for mental disorders.

In psychiatry, polypharmacy is defined as a patient with at least two concurrent psychiatric medications [@Shrivastava2013]. With a prevalence between 13-90%, psychiatric polypharmacy manifests in five categories, namely same-class, multi-class, adjunctive, augmentation, and total polypharmacy. Same-class polypharmacy is the use of multiple medications from the same class. Multi-class polypharmacy is the use of multiple medications from different classes indicated for the same symptom cluster. Adjunctive polypharmacy is the use of additional medications to treat side effects due to other medications. Augmentation polypharmacy is the use of full-dose and sub-dose medications from a different class for the same symptom cluster. Finally, total polypharmacy is the total count of all medications used by a patient, both psychopharmaca and other classes alike [@Shrivastava2013]. For policymakers, differences in defining polypharmacy pose a challenge to design a monitoring system [@Sirois2016].

Monitoring polypharmacy regimens requires a holistic approach since drug prescription highly depends on the physician's prescribing habits [@Shrivastava2013], patient-physician relationship, and pharmaceutical market, which all bear an interplay with the decision-making process [@Cavallo2012]. A holistic polypharmacy monitoring system needs to describe the importance of polypharmacy by incorporating both qualitative and quantitative measures [@Sirois2016]. Quantitatively attributing importance to each medication is challenging, accounting for polypharmacy as a complex and interlinked medication uses. To reduce the complexity, graph theory offers a complementary approach by analyzing polypharmacy as a DPN. Transforming prescription registry into a DPN has been previously investigated, where DPN-based approach represented the general trend in co-prescription [@Cavallo2012; @Bazzoni2015].

A DPN is constructed as a matrix which rows to columns represent pairs of medication being taken by a patient in a day. Since there is no limit for the number of rows and columns, a DPN can contain virtually every possible combination of polypharmacy regimens. A DPN is generalizable as a population aggregate by coalescing all single-patient DPN through matrix addition [@Cavallo2012]. Constructing a DPN allows examining co-prescription dynamics in the population as graph measures, which is also attributable to each medication. One of such measures is graph centrality, where we calculate scores representing the number of co-medications, number of connecting paths, average distance, and relative importance [@Askar2021]. The relative importance score is of particular interest to address concerns raised by @Sirois2016 and @Delara2022 regarding standardized quantitative indicator of polypharmacy. Relative importance, most commonly measured by eigenvector centrality, represents the connection to vital medications. A high eigenvector centrality in a DPN indicates more prescriptions with highly-prescribed medications. Following up @Cavallo2012 and @Bazzoni2015, we hypothesize that eigenvector centrality as a measure of relative importance is attributable to indicate polypharmacy. Our investigation showcases eigenvector centrality as a quantitative indicator of polypharmacy to promote a systematic public health monitoring system.
